My title

### 2024-10-30 paper/paper.qmd
test of model

library(tidyverse)
library(dplyr)

data <- read.csv("../data/01-raw_data/raw_data.csv")
```

```{r}
# pollsters measured by highest numeric_grade, transparency_score, pollster_rating_id

usdata <- data %>%
  group_by(pollster) %>%
  summarize(
    pollster_rating_id = mean(pollster_rating_id, na.rm = TRUE),
    numeric_grade = mean(numeric_grade, na.rm = TRUE),
    pollscore = mean(pollscore, na.rm = TRUE),
    transparency_score = mean(transparency_score, na.rm = TRUE)
  ) %>%
  filter(numeric_grade >= 1.5, transparency_score >= 5, pollster_rating_id >= 400) %>%
    arrange(desc(pollster_rating_id))
usdata

# pollscore is negative ? how is it measured 
```
# Introduction

@@

>>>>

library(tidyverse)
library(dplyr)
library(janitor)
```

#Load and clean data
data <- read.csv("../data/01-raw_data/raw_data.csv") |>
  clean_names()
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Filter NYT polls and remove entries with NA or empty value for statr
# Select imporntat variables (change as you please)

preddata <- data %>%
 # filter(pollster_id == 1424 & !is.na(state) & state != "") %>%
  select(display_name, question_id, sample_size, state, party, answer, candidate_id, candidate_name, pct, start_date, end_date)

head(preddata)
```

```{r}
#| echo: false
#| warning: false
#| message: false

#FILTERED SWING STATES, SUBJECT TO CHANGE, ALSO INCLUDED NATIONAL?

harrisdata <- preddata %>%
  filter(state %in% c("Georgia", "Arizona", "Nevada", "Michigan", "North Carolina", "Pennsylvania", "")) %>%
  filter(candidate_name == "Kamala Harris") %>%
  mutate(
    state = if_else(state == "", "National", state),
    end_date = mdy(end_date)
  ) %>%

  # Step 4: Filter by date and calculate number of respondents
  filter(end_date >= as.Date("2024-07-21")) %>%
  mutate(
    num_harris = round((pct / 100) * sample_size, 0)
  )

# Plot the filtered and transformed data
base_plot <- ggplot(harrisdata, aes(x = end_date, y = pct)) +
  theme_classic() +
  labs(y = "Harris percent", x = "Date") +
  geom_point(aes(color = state)) +
  geom_smooth() +
  theme(legend.position = "bottom")

base_plot

base_plot +
  geom_point() +
  geom_smooth()

```
```{r}
#| echo: false
#| warning: false
#| message: false

## MODEL 1, LINEAR MODEL WITH PREDICTOR END_DATE

model_date <- lm(pct ~ end_date, data = harrisdata)

harrisdata <- harrisdata |>
  mutate(
    fitted_date = predict(model_date)
  )

ggplot(harrisdata, aes(x = end_date)) +
  geom_point(aes(y = pct), color = "black") +
  geom_line(aes(y = fitted_date), color = "blue", linetype = "dotted") +
  theme_classic() +
  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date")
```

## MODEL 2, BAYESIAN MODEL WITH PREDICTOR END_DATE

harrisdata <- harrisdata |>
  mutate(state = factor(state))

model1 <- cbind(num_harris, sample_size - num_harris) ~ (1 | state)

priors <- normal(0, 2.5, autoscale = TRUE)

bayesian_model_1 <- stan_glmer(
  formula = model1,
  data = harrisdata,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 4,
  adapt_delta = 0.95
)
```

```{r}
#| echo: false
#| warning: false
#| message: false

pp_check(bayesian_model_1)

summary(bayesian_model_1)

plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)

```

```{r}
harrisdata <- harrisdata |>
  mutate(
    end_date_num = as.numeric(end_date - min(end_date))
  )

spline_model <- stan_glm(
  pct ~ ns(end_date_num, df = 6),
  data = harrisdata,
  family = gaussian(),
  prior = normal(0, 5),
  prior_intercept = normal(50, 10),
  seed = 1234,
  iter = 2000,
  chains = 4,
  refresh = 0
)

summary(spline_model)
pp_check(spline_model)

# Summarize the model
summary(spline_model)

# Posterior predictive checks
pp_check(spline_model)
```

```{r}
new_data <- data.frame(
  end_date_num = seq(
    min(harrisdata$end_date_num),
    max(harrisdata$end_date_num),
    length.out = 100))

posterior_preds <- posterior_predict(spline_model, newdata = new_data)

pred_summary <- new_data |>
  mutate(
    pred_mean = colMeans(posterior_preds),
    pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975)
  )

ggplot(harrisdata, aes(x = end_date_num, y = pct)) +
  geom_point() +
  geom_line(
    data = pred_summary,
    aes(x = end_date_num, y = pred_mean),
    color = "blue",
    inherit.aes = FALSE
  ) +
  geom_ribbon(
    data = pred_summary,
    aes(x = end_date_num, ymin = pred_lower, ymax = pred_upper),
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  labs(
    x = "Days since earliest poll",
    y = "Percentage",
    title = "Poll Percentage over Time with Spline Fit"
  ) +
  theme_minimal()

```

# Introduction

@@

Why we choose NYT?
- `numeric_grade` is 3
- `pollscore` is -1.5, a score of reliability called "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically.", where negative numbers are better
- `transparency_score` is 9, reflects pollsters transparency about their methodology (calculated based on how much information it discloses about its polls and weighted by recency)
- `population_full` is 'rv', respondents are registered voters
